---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# quarrisk

<!-- badges: start -->
[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](LICENSE)
<!-- badges: end -->

## Overview

**quarrisk** provides an interactive Shiny dashboard and helper functions to explore **quarantine outbreak risk** under different assumptions about basic reproduction number (**R₀**), vaccine effectiveness (**VE**), and vaccination **coverage**.

It ships three tidy datasets:

- `ob_probs_travellers` — outbreak probabilities by traveller/worker pathway  
- `ob_timings_summary` — median (50%) and tail (95%) days until outbreak  
- `ob_summary` — a joined grid of probabilities + timings across R₀ × VE × coverage

The dashboard has three tabs: **Overview**, **Risk Explorer**, and **Timing Summary**.

## Features

- 📦 **Three curated datasets** covering probabilities and timing  
- 🧭 **Overview**: nearest-scenario readout + quick comparison of traveller vs worker risk  
- 📈 **Risk Explorer**: lines and heatmaps to see how risk changes with VE, R₀, and coverage  
- ⏱️ **Timing Summary**: median and tail time-to-outbreak profiles  
- 🧰 **Helpers**: `slice_scenario()` to grab scenarios; `plot_risk_vs_ve()` to plot VE curves  
- 🖥️ **One-line app launcher**: `run_app()`

## Installation

You can install the development version of quarrisk from [GitHub](https://github.com/) with:

```{r installation-guide, eval=FALSE}

install.packages("remotes")
remotes::install_github("ETC5523-2025/assignment-4-packages-and-shiny-apps-Yashitta")

```

## Quick Start

```{r load libraries}

library(quarrisk)

```

## Explore the datasets

```{r example, eval=FALSE}

data(ob_probs_travellers)
head(ob_probs_travellers)
data(ob_timings_summary)
head(ob_timings_summary)
data(ob_summary)
dplyr::glimpse(ob_summary)

```

### Launch interactive dashboard

```{r launch-run-app, eval=FALSE}

run_app()

```

## Dashboard Views

- **Overview** – Shows the nearest scenario to your chosen **R₀ / VE / coverage**, value boxes, and a quick traveller vs worker probability comparison.

- **Risk Explorer** –
  - **Lines:** outbreak probability vs **VE** at fixed **R₀** & **coverage**
  - **Heatmap:** probability across **R₀ × coverage** at fixed **VE**

- **Timing Summary** – Median (**chance50**) and tail (**chance95**) days vs **coverage / R₀** for the selected **VE**.

## Example Analysis

Compute the gap between traveller and worker outbreak probabilities for a given scenario grid:

```{r gap-table, echo=FALSE, message=FALSE, warning=FALSE}

library(dplyr)
library(knitr)
library(scales)

tab <- ob_summary |>
  mutate(
    gap = traveller_ob_prob - worker_ob_prob,
    across(c(traveller_ob_prob, 
             worker_ob_prob, gap), 
           ~ percent(.x, accuracy = 0.1))) |>
  arrange(desc(abs(readr::parse_number(gap)))) |>
  select(r0, ve, coverage, traveller_ob_prob, worker_ob_prob, gap) |>
  head(6)

kable(
  tab,
  format = "pipe",    # GitHub-friendly markdown table
  col.names = c("R₀", "VE", "Coverage", "Traveller prob", "Worker prob", "Gap"),
  align = "rrrccc")

```

## Available Datasets

- `ob_probs_travellers` — Tibble with columns: `r0`, `ve`, `coverage`, `traveller_ob_prob`, `worker_ob_prob`.
- `ob_timings_summary` — Tibble with columns: `r0`, `ve`, `coverage`, `chance50`, `chance95`.
- `ob_summary` — Join of the two above across the full scenario grid.

## Functions

- `run_app()` — Launch the Shiny dashboard.
- `slice_scenario(r0, ve, coverage)` — Return rows matching (or nearest to) a scenario.
- `plot_risk_vs_ve(r0, coverage, pathway = c("traveller_ob_prob","worker_ob_prob"))` — Line plot of outbreak probability vs VE at fixed **R₀** & coverage.

## Documentation

- Package repo: <https://github.com/ETC5523-2025/assignment-4-packages-and-shiny-apps-Yashitta.git>  
